```{r}
install.packages("forcats")
library(forcats)
install.packages("dplyr")
library(dplyr)
install.packages("stringr")
library(stringr)
```

```{r}
review=read.csv("review.csv")
review$id=NULL
review$X=NULL
review$review_scores_rating=as.factor(review$review_scores_rating)
review=na.omit(review)
```


## Training Set EDA
```{r}
par(mfrow=c(2,3))
hist(review$host_listings_count, col="blue")
plot(review$cancellation_policy, col="blue")
plot(review$host_identity_verified, col="blue")
plot(review$host_is_superhost, col="blue")
plot(review$host_response_time, col="blue")
plot(review$room_type, col="blue")
plot(review$bed_type, col="blue")
plot(review$instant_bookable, col="blue")
plot(review$property_type, col="blue")

plot(as.factor(review$bedrooms), col="blue")
plot(as.factor(review$bathrooms), col="blue")
plot(as.factor(review$minimum_nights), col="blue")
plot(as.factor(review$maximum_nights), col="blue")
plot(as.factor(review$guests_included), col="blue")
plot(as.factor(review$accommodates), col="blue")

hist(review$host_listings_count, col="blue")
plot(review$property_type, col="blue")
plot(review$bed_type, col="blue")
#?? what about the other variables
hist(as.numeric(review$price))
hist(as.numeric(review$cleaning_fee))

ct=table(review$review_scores_rating)
prop.ct=prop.table(ct)
```
```{r}
plot(review$property_type, review$review_scores_rating)
plot(review$bed_type, review$review_scores_rating)
```


## Data Preprocessing
```{r}
review$property_type = fct_recode(review$property_type,
                                  "House/Apartment"="House",
                                  "House/Apartment"="Apartment",
                                  "Other"="Condominium",
                                  "Other"="Townhouse",
                                  "Other"="Tent",
                                  "Other"="Guest suite",
                                  "Other"="Serviced apartment",
                                  "Other"="Cottage",
                                  "Other"="Camper/RV",
                                  "Other"="Bungalow",
                                  "Other"="Loft",
                                  "Other"="Other",
                                  "Other"="Boat",
                                  "Other"="Bed and breakfast",
                                  "Other"="Guesthouse",
                                  "Other"="In-law",
                                  "Other"="Farm stay",
                                  "Other"="Villa",
                                  "Other"="Boutique hotel",
                                  "Other"="Cabin",
                                  "Other"="Yurt",
                                  "Other"="Tiny house",
                                  "Other"="Aparthotel",
                                  "Other"="Hostel",
                                  "Other"="Houseboat",
                                  "Other"="Resort",
                                  "Other"="Treehouse")

review$bed_type = fct_recode(review$bed_type,
                             "Real Bed"="Real Bed",
                             "Other"="Futon",
                             "Other"="Airbed",
                             "Other"="Couch",
                             "Other"="Pull-out Sofa")

review$host_listings_count = ifelse(review$host_listings_count==1, "Single", "Multiple")

review$cancellation_policy = fct_recode(review$cancellation_policy,
                                         "Flexible" = 'flexible',
                                         "Moderate" = 'moderate',
                                         "Strict" = 'strict_14_with_grace_period',
                                         "Strict" = 'strict',
                                         "Super Strict" = 'super_strict_60',
                                         "Super Strict" = 'super_strict_30')

review$host_response_rate = as.numeric(review$host_response_rate)
review$price = as.numeric(review$price)
review$cleaning_fee = as.numeric(review$cleaning_fee)

neighborhoods.centers <- as.data.frame(matrix(0, nrow = 17, ncol = 2))
rownames(neighborhoods.centers) <- unique(review$neighbourhood_group_cleansed)
colnames(neighborhoods.centers) <- c('latitude', 'longitude')
neighborhoods.centers['Other neighborhoods',1:2] <- c(47.6097, -122.3422)
neighborhoods.centers['Lake City',1:2] <- c(47.723244, -122.286596)
neighborhoods.centers['Queen Anne',1:2] <- c(47.6205, -122.3493)
neighborhoods.centers['University District',1:2] <- c(47.655548, -122.303200)
neighborhoods.centers['Delridge',1:2] <- c(47.560966, -122.351856)
neighborhoods.centers['Capitol Hill',1:2] <- c(47.6253, -122.3222)
neighborhoods.centers['Downtown',1:2] <- c(47.6097, -122.3422)
neighborhoods.centers['Seward Park',1:2] <- c(47.5496, -122.2574)
neighborhoods.centers['Ballard',1:2] <- c(47.6655, -122.3970)
neighborhoods.centers['West Seattle',1:2] <- c(47.532411, -122.393295)
neighborhoods.centers['Central Area',1:2] <- c(47.5932, -122.3005)
neighborhoods.centers['Cascade',1:2] <- c(47.6205, -122.3493)
neighborhoods.centers['Beacon Hill',1:2] <- c(47.579265, -122.311722)
neighborhoods.centers['Rainier Valley',1:2] <- c(47.529612, -122.263677)
neighborhoods.centers['Magnolia',1:2] <- c(47.658198, -122.405829)
neighborhoods.centers['Northgate',1:2] <- c(47.706164, -122.325904)
neighborhoods.centers['Interbay',1:2] <- c(47.643620, -122.378368)

neighborhood <- as.character(review$neighbourhood_group_cleansed)
distance <- sqrt((review$latitude - neighborhoods.centers[neighborhood,1])^2 +
                   (review$longitude - neighborhoods.centers[neighborhood,2])^2)
miles <- distance * 69
review$distance = cut(miles,
            breaks = c(0, 1, 2, Inf),
            labels = c('center', 'close', 'far'))

review$latitude <- NULL
review$longitude <- NULL

review$amenities_count <- str_count(review$amenities, ',') + 1
review$amenities <- NULL
```


## Model 1: Logistic Regression (redo training/test error rates and consider omitting variables)
```{r}
set.seed(1)
train=sample(1:nrow(review), floor(0.3*(nrow(review))), replace=FALSE) # split 30% to be used as validation set
review.test=review[-train,]
review.train=review[train,]

model1 = glm(review_scores_rating~., family = "binomial", data=review.train)
summary(model1)$coef

glm.probs=predict(model1, newdata=review.test, type="response")
glm.pred=rep(0, nrow(review))
glm.pred[glm.probs>0.5]=1

mean(glm.pred==review.train$review_scores_rating)
# test set error rate is 21.12%
```

```{r}
#library(MASS)
#lda.fit = lda(review_scores_rating~., data=review, subset=train)
#lda.pred = predict(lda.fit, newdata=review[-train,])
```


## Model 2: Classification Tree
```{r}
install.packages("tree")
library(tree)

tree.review=tree(review_scores_rating~., data=review)
summary(tree.review)

# training error rate is 12.59%

tree.train=tree(review_scores_rating~., data=review.train)
tree.pred=predict(tree.train, newdata=review.test, type="class")
table(tree.pred, review.test$review_scores_rating)
(113+2327)/nrow(review.test)

# 86.19% of the test observations are correctly classified
```
```{r}
cv.review=cv.tree(tree.train, FUN=prune.misclass)
prune.review=prune.misclass(tree.train, best=5)
review.pred=predict(prune.review, newdata=review.test, type="class")
table(review.pred, review.test$review_scores_rating)

(85+2362)/nrow(review.test)

# 86.44% of the test observations are correctly classified
```

## Model 3: K-NN ?? standardize
```{r}
library(class)
# one-hot encoding 
dummy=model.matrix(~.,data=review.train)
dummy=as.data.frame(dummy)
dummy$`(Intercept)`=NULL
train.score=dummy$review_scores_rating1

dummy.test=model.matrix(~., data=review.test)
dummy.test=as.data.frame(dummy.test)
dummy.test$`(Intercept)`=NULL
test.score = dummy.test$review_scores_rating1

knn.pred = knn(dummy, dummy.test, train.score, k=1)
table(knn.pred, test.score)
(143+2194)/nrow(dummy.test)

# 82.55% of the test observations are correctly classified
```


## Model 4: SVM ?? standardize // radial, polynomial, linear // cost, degree, gamma
```{r}
install.packages("e1071")
library(e1071)
# wrong method because categorical variables not encoded (!!!)

#svmfit = svm(review_scores_rating~., data=review.train, kernel="polynomial", cost=1e7)
#summary(svmfit)

tune.out = tune(svm, review_scores_rating~., data=review.train, kernel="linear", ranges=list(cost=c(0.001, 0.01, 1, 10, 100)))
summary(tune.out)

bestmod=tune.out$best.model

ypred=predict(bestmod, newdata=review.test)
table(predict=ypred, truth=review.test$review_scores_rating)
(134+2320)/nrow(review.test)

# 86.19% of the test observations are correctly classified
```

```{r}
#svmfit=svm(review_scores_rating1~., data=dummy, kernel="polynomial", cost=1e7, scale=TRUE)

tune.out = tune(svm, review_scores_rating~., data=dummy, kernel="polynomial", scale=TRUE, ranges=list(cost=c(0.001, 0.01, 1, 10, 100), gamma=c(0.01, 10, 100)))
summary(tune.out)

bestmod=tune.out$best.model

#ypred=predict(bestmod, newdata=dummy.test)
#table(predict=ypred, truth=dummy.test$review_scores_rating1)

#plot(bestmod, dummy)
```

```{r}
tune.out = tune(svm, review_scores_rating~., data=dummy, kernel="radial", scale=TRUE, ranges=list(cost=c(0.001, 0.01, 1, 10, 100), gamma=c(0.01, 10, 100)))
summary(tune.out)

bestmod=tune.out$best.model
```


## Model 5: Random Forest (must preprocess data to numeric values)
```{r}
#install.packages("randomForest")
#library(randomForest)

#bag.review=randomForest(review_scores_rating~., data=review.train, nodesize=10, mtry=20, ntree=50)
review.train=review[train,]
bag.review=randomForest(review_scores_rating~., data=review.train)
bag.review

which.min(bag.review$err.rate[,1]) # 274 trees is plenty
#bag.review=randomForest(review_scores_rating~., data=review.train, mtry=20)

plot(bag.review, ylim=c(0.11, 0.17), main="Untuned Model Error")
plot(bag.review$err.rate[,1], pch=19)
```

```{r}
features <- setdiff(names(review), "review_scores_rating")

rf2 <- tuneRF(
  x          = review.train[features],
  y          = review.train$review_scores_rating,
  ntreeTry   = 280,
  mtryStart  = 5,
  stepFactor = 1.5,
  improve    = 0.01,
  trace      = FALSE      # to not show real-time progress 
)
```

```{r}
# !!! good model !!!
new.bag=randomForest(review_scores_rating~., data=review.train, ntree=270, mtry=20, nodesize=9)
plot(new.bag, ylim=c(0.117, 0.2), main="Tuned Model Error")
```
```{r}
reduced.mod = randomForest(review_scores_rating~.-, data=review.train, ntree=270, mtry=20, nodesize=9)
```

```{r}
yhat.bag=predict(new.bag, newdata=review.test)
#yhat.bag=predict(bag.review, newdata=review.test)
1-(sum(diag(table(yhat.bag, review.test$review_scores_rating)))/nrow(review.test))
```

```{r}
install.packages("ranger")
library(ranger)

hyper_grid <- expand.grid(
  mtry       = seq(10, 20, by = 2),
  node_size  = seq(3, 9, by = 2)
)

for(i in 1:nrow(hyper_grid)) {
  # train model
  model <- ranger(
    formula         = review_scores_rating~., 
    data            = review.train, 
    num.trees       = 270,
    mtry            = hyper_grid$mtry[i],
    min.node.size   = hyper_grid$node_size[i]
  )
}

model
```


## Model 6: Boosting ?? tuning parameters
```{r}
install.packages("gbm")
library(gbm)

review.train2=review[train,]

boost.review=gbm(review_scores_rating~., data=review.train2, distribution="bernoulli", n.trees=5000, shrinkage = 0.001, cv.folds = 5)

summary(boost.review)
boost.review

review.test2=review[-train,]

yhat.boost=predict(boost.review, newdata=review.test2, n.trees=5000)
yhat.boost
mean((as.numeric(yhat.boost)-as.numeric(review.test2$review_scores_rating))^2)
```


## Unsupervised Learning EDA ??

## PCA does not work on categorical
```{r}
# PCA
mat = data.matrix(review.train2, rownames.force = NA)
pr = prcomp(mat, center=TRUE, scale=TRUE)
plot(pr$x[,1:2], pch=19)

summary(pr)

pve = 100*pr$sdev^2/sum(pr$sdev^2)
plot(pve, type="o", col="blue")
plot(cumsum(pve), type="o", col="brown3")
```

```{r}
# hierarchical clustering ??
review=na.omit(review)
review$longitude=NULL
review$latitude=NULL
review$amenities=NULL

d = dist(review, method="euclidean")
hc1 = hclust(d, method="complete")
hc2 = hclust(d, method="average")
hc3 = hclust(d, method="single")

cut1=cutree(hc1, k=2)
cut2=cutree(hc2, k=2)
cut3=cutree(hc3, k=2)
plot(review$review_scores_rating, col=factor(cut1))
plot(review$review_scores_rating, col=factor(cut2))
plot(review$review_scores_rating, col=factor(cut3))

plot(hc1, cex=0.6)
plot(hc2, cex=0.6)
plot(hc3, cex=0.6)
```

```{r}
# more categorical preprocessing for svm/random forest/boosting
review$cancellation_policy <- fct_recode(review$cancellation_policy,
                                         "1" = 'Flexible',
                                         "2" = 'Moderate',
                                         "3" = 'Strict',
                                         "4" = 'Super Strict')

review$distance <- fct_recode(review$distance,
                                         "1" = 'close',
                                         "2" = 'center',
                                         "3" = 'far')

review$bed_type <- fct_recode(review$bed_type,
                              "3" = "Couch",
                              "2" = "Airbed", "2" = "Futon", "2" = "Pull-out Sofa",
                              "1" = "Real Bed")
levels(review$host_is_superhost) = c(0,1)
levels(review$host_identity_verified) = c(0,1)
levels(review$instant_bookable) = c(0,1)

review$host_response_time = fct_recode(review$host_response_time,
                                         "1" = "within an hour",
                                         "2" = "within a few hours",
                                         "3" = "within a day",
                                         "4" = "a few days or more")

review$property_type = fct_recode(review$property_type,
                                  "1"="House/Apartment",
                                  "2"="Condominium",
                                  "2"="Townhouse",
                                  "2"="Tent",
                                  "2"="Guest suite",
                                  "2"="Serviced apartment",
                                  "2"="Cottage",
                                  "2"="Camper/RV",
                                  "2"="Bungalow",
                                  "2"="Loft",
                                  "2"="Other",
                                  "2"="Boat",
                                  "2"="Bed and breakfast",
                                  "2"="Guesthouse",
                                  "2"="In-law",
                                  "2"="Farm stay",
                                  "2"="Villa",
                                  "2"="Boutique hotel",
                                  "2"="Cabin",
                                  "2"="Yurt",
                                  "2"="Tiny house",
                                  "2"="Aparthotel",
                                  "2"="Hostel",
                                  "2"="Houseboat",
                                  "2"="Treehouse",
                                  "2"="Resort")

review$room_type = fct_recode(review$room_type,
                              "1"="Private room",
                              "2"="Entire home/apt",
                              "3"="Shared room")

review$host_listings_count = ifelse(review$host_listings_count==1, 0, 1)
```

## EM does not work on categorical
```{r}
install.packages("mixtools")
library(mixtools)

mix=mvnormalmixEM(dummy, k=2)
dummy.mat=as.matrix(dummy)
```

## K-Means N/A on categorical 
```{r}
km.out=kmeans(dummy, centers=2, nstart=50, algorithm="Lloyd")

tab=xtabs(~km.out$cluster+dummy$review_scores_rating1)
tab

1-sum(diag(tab))/sum(tab)
```

## PCAmix for both categorical/numerical mix data
```{r}
install.packages("PCAmixdata")
library(PCAmixdata)

split=splitmix(review.train)
x1=split$X.quanti
x2=split$X.quali

res.pcamix=PCAmix(X.quanti=x1, X.quali=x2,rename.level=TRUE, graph=FALSE)
summary(res.pcamix)

plot(res.pcamix, coloring.ind = x2$review_scores_rating)

plot(res.pcamix$eig[,1]) # retain 11 components
res.pcamix$eig # 13 dimensions needed to explain 50% of variance
```

```{r}
test.set=read.csv("review_test.csv")
test.set$id=NULL
test.set$X=NULL

test.set$property_type = fct_recode(test.set$property_type,
                                  "House/Apartment"="House",
                                  "House/Apartment"="Apartment",
                                  "Other"="Condominium",
                                  "Other"="Townhouse",
                                  "Other"="Tent",
                                  "Other"="Guest suite",
                                  "Other"="Serviced apartment",
                                  "Other"="Cottage",
                                  "Other"="Camper/RV",
                                  "Other"="Bungalow",
                                  "Other"="Loft",
                                  "Other"="Other",
                                  "Other"="Boat",
                                  "Other"="Bed and breakfast",
                                  "Other"="Guesthouse",
                                  "Other"="In-law",
                                  "Other"="Farm stay",
                                  "Other"="Villa",
                                  "Other"="Boutique hotel",
                                  "Other"="Cabin",
                                  "Other"="Yurt",
                                  "Other"="Tiny house",
                                  "Other"="Aparthotel",
                                  "Other"="Hostel",
                                  "Other"="Houseboat",
                                  "Other"="Resort",
                                  "Other"="Treehouse")

test.set$bed_type = fct_recode(test.set$bed_type,
                             "Real Bed"="Real Bed",
                             "Other"="Futon",
                             "Other"="Airbed",
                             "Other"="Couch",
                             "Other"="Pull-out Sofa")

test.set$host_listings_count = ifelse(test.set$host_listings_count==1, "Single", "Multiple")

test.set$cancellation_policy = fct_recode(test.set$cancellation_policy,
                                         "Flexible" = 'flexible',
                                         "Moderate" = 'moderate',
                                         "Strict" = 'strict_14_with_grace_period',
                                         "Strict" = 'strict',
                                         "Super Strict" = 'super_strict_60',
                                         "Super Strict" = 'super_strict_30')

test.set$host_response_rate = as.numeric(test.set$host_response_rate)
test.set$price = as.numeric(test.set$price)
test.set$cleaning_fee = as.numeric(test.set$cleaning_fee)

neighborhoods.centers <- as.data.frame(matrix(0, nrow = 17, ncol = 2))
rownames(neighborhoods.centers) <- unique(test.set$neighbourhood_group_cleansed)
colnames(neighborhoods.centers) <- c('latitude', 'longitude')
neighborhoods.centers['Other neighborhoods',1:2] <- c(47.6097, -122.3422)
neighborhoods.centers['Lake City',1:2] <- c(47.723244, -122.286596)
neighborhoods.centers['Queen Anne',1:2] <- c(47.6205, -122.3493)
neighborhoods.centers['University District',1:2] <- c(47.655548, -122.303200)
neighborhoods.centers['Delridge',1:2] <- c(47.560966, -122.351856)
neighborhoods.centers['Capitol Hill',1:2] <- c(47.6253, -122.3222)
neighborhoods.centers['Downtown',1:2] <- c(47.6097, -122.3422)
neighborhoods.centers['Seward Park',1:2] <- c(47.5496, -122.2574)
neighborhoods.centers['Ballard',1:2] <- c(47.6655, -122.3970)
neighborhoods.centers['West Seattle',1:2] <- c(47.532411, -122.393295)
neighborhoods.centers['Central Area',1:2] <- c(47.5932, -122.3005)
neighborhoods.centers['Cascade',1:2] <- c(47.6205, -122.3493)
neighborhoods.centers['Beacon Hill',1:2] <- c(47.579265, -122.311722)
neighborhoods.centers['Rainier Valley',1:2] <- c(47.529612, -122.263677)
neighborhoods.centers['Magnolia',1:2] <- c(47.658198, -122.405829)
neighborhoods.centers['Northgate',1:2] <- c(47.706164, -122.325904)
neighborhoods.centers['Interbay',1:2] <- c(47.643620, -122.378368)

neighborhood <- as.character(test.set$neighbourhood_group_cleansed)
distance <- sqrt((test.set$latitude - neighborhoods.centers[neighborhood,1])^2 +
                   (test.set$longitude - neighborhoods.centers[neighborhood,2])^2)
miles <- distance * 69
test.set$distance = cut(miles,
            breaks = c(0, 1, 2, Inf),
            labels = c('center', 'close', 'far'))

test.set$latitude <- NULL
test.set$longitude <- NULL

test.set$amenities_count <- str_count(test.set$amenities, ',') + 1
test.set$amenities <- NULL

test.set$cancellation_policy <- fct_recode(test.set$cancellation_policy,
                                         "1" = 'Flexible',
                                         "2" = 'Moderate',
                                         "3" = 'Strict',
                                         "4" = 'Super Strict')

test.set$distance <- fct_recode(test.set$distance,
                                         "1" = 'close',
                                         "2" = 'center',
                                         "3" = 'far')

test.set$bed_type <- fct_recode(test.set$bed_type,
                              "3" = "Couch",
                              "2" = "Airbed", "2" = "Futon", "2" = "Pull-out Sofa",
                              "1" = "Real Bed")
levels(test.set$host_is_superhost) = c(0,1)
levels(test.set$host_identity_verified) = c(0,1)
levels(test.set$instant_bookable) = c(0,1)

test.set$host_response_time = fct_recode(test.set$host_response_time,
                                         "1" = "within an hour",
                                         "2" = "within a few hours",
                                         "3" = "within a day",
                                         "4" = "a few days or more")

test.set$property_type = fct_recode(test.set$property_type,
                                  "1"="House/Apartment",
                                  "2"="Condominium",
                                  "2"="Townhouse",
                                  "2"="Tent",
                                  "2"="Guest suite",
                                  "2"="Serviced apartment",
                                  "2"="Cottage",
                                  "2"="Camper/RV",
                                  "2"="Bungalow",
                                  "2"="Loft",
                                  "2"="Other",
                                  "2"="Boat",
                                  "2"="Bed and breakfast",
                                  "2"="Guesthouse",
                                  "2"="In-law",
                                  "2"="Farm stay",
                                  "2"="Villa",
                                  "2"="Boutique hotel",
                                  "2"="Cabin",
                                  "2"="Yurt",
                                  "2"="Tiny house",
                                  "2"="Aparthotel",
                                  "2"="Hostel",
                                  "2"="Houseboat",
                                  "2"="Treehouse",
                                  "2"="Resort")

test.set$room_type = fct_recode(test.set$room_type,
                              "1"="Private room",
                              "2"="Entire home/apt",
                              "3"="Shared room")

test.set$host_listings_count = ifelse(test.set$host_listings_count==1, 0, 1)
```

```{r}
test.bag=predict(new.bag, newdata=test.set)

review.guesses=as.vector(test.bag)

review.acc=sum(diag(table(yhat.bag, review.test$review_scores_rating)))/nrow(review.test)
```

```{r}
save(list=c("review.guesses","review.acc"),file="stat462final.RData")
price.guesses
price.mse
team.name="beast"

save(list=c("review.guesses","review.acc","price.guesses","price.mse", "team.name"), file="stat462final.RData")
```

